---
title: "qPCR-analysis-for-manuscript"
author: "diana baetscher"
date: "2023-11-28"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-libraries}
library(readxl)
library(patchwork)
library(tidyverse)
library(ggpubr)
```


## Clean analyses with the new qPCR data for manuscript

28 November 2023

```{r load-data}
# read in the extraction database metadata
metadata <- read_xlsx("../data/extraction_database.xlsx", sheet = "metadata")

# read in the combined qPCR results
#qpcr <- read_xlsx("../data/qPCR_results_combined.xlsx")

# organize color palette
mycols <- c("coral3", "cadetblue4")
```


### Metadata QC

```{r meta-qc}
# clean up the metadata format and variable names
frsh_meta <- metadata %>%
  mutate(tide = ifelse(`Sample Code` == "AM", "incoming", "outgoing")) %>%
  mutate(tide = ifelse(is.na(`Sample Code`) | `Sample Code` == "Creek", NA, tide)) %>%
  rename(sample = `Extraction ID`, location = `Sample Code`, label = `Sample Label`, type = `Sample_type`) %>%
  # distance needs to be a numeric variable without meters attached
  mutate(dist_grp = ifelse(location %in% c("PM", "AM"), label, NA))

# strip off the meters on dist_grp so we can order them properly
frsh_meta$dist_grp <- gsub("[^0-9.-]", "", frsh_meta$dist_grp)

# make another variable that is numeric distance (not character)
clean_meta <- frsh_meta %>%
  mutate(distance = as.numeric(dist_grp)) %>%
  # remove some of the unneeded variables
  dplyr::select(sample, location, label, type, tide, dist_grp, distance) %>%
  # add info to the creek samples
  mutate(label = ifelse(location == "Creek", "freshwater", label)) %>%
  mutate(label = ifelse(is.na(location), "blank", label))


clean_meta
```


### qPCR data

## Combine data

Patterns across samples of inhibition?? Look at metadata.

Read in all combined data:
```{r}
file_path <- "../data/qpcr/"

file_path %>%
  list.files() %>%
  .[str_detect(., ".xlsx")] -> xlsx_file_names

# Load everything into the Global Environment
xlsx_file_names %>%
  purrr::map(function(file_name){ # iterate through each file name
  
  read_xlsx(paste0(file_path, file_name), sheet = "Results", skip = 34)
  
}) -> df_list_read2 # Assign to a list

# combine all files into one df
# nothing was done to the quantities in this df
all_plates2021 <- bind_rows(df_list_read2, .id = "plate") %>%
    mutate(efficiency = -1 + 10^ (-1/Slope))

# calculate efficiency for all plates
all_plates2021 %>%
  group_by(`Target Name`) %>%
  summarise(E = mean(efficiency))

```

```{r combine-qPCR-and-metadata}
# but I'm going to want to add a column that designates whether the sample is inhibited or not.
qpcr_data_w_metadata <- all_plates2021 %>%
  left_join(., clean_meta, by = c("Sample Name" = "sample"))

qpcr_data_w_metadata %>%
  filter(Task == "UNKNOWN" &
           CT != "Undetermined") %>%
  dplyr::select(`Sample Name`, `Ct Mean`, dist_grp, tide, distance) %>%
  unique()

```

Similar to the undiluted plates, I'll want to read in all three results documents and then analyze them in tandem.

### Diluted samples from 2021

```{r read-in-list-of-inhibited-sample-results}
diluted_file_path <- "../data/diluted_qpcr/"

diluted_file_path %>%
  list.files() %>%
  .[str_detect(., ".xlsx")] -> xlsx_file_names_diluted

# Load everything into the Global Environment
xlsx_file_names_diluted %>%
  purrr::map(function(file_name){ # iterate through each file name
  
  read_xlsx(paste0(diluted_file_path, file_name), sheet = "Results", skip = 34)
  
}) -> df_list_read3 # Assign to a list

diluted_plates <- bind_rows(df_list_read3, .id = "plate")

# multiple quantities by 10 because of 1:10 dilution factor
diluted_plates_10x <- diluted_plates %>%
  mutate(quant = Quantity*10) %>%
  mutate(inhibited = "yes")

# list of inhibited samples
inhibited_list <- diluted_plates_10x %>%
  filter(Task =="UNKNOWN") %>%
  dplyr::select(`Sample Name`) %>%
  unique()

```




### Combined data from inhibited samples and uninhibited samples

```{r}
combined_data <- all_plates2021 %>%
  anti_join(., inhibited_list) %>%
  bind_rows(., diluted_plates_10x) %>%
  filter(Task == "UNKNOWN" & 
           `Target Name` == "Oke_COI") %>%
  #select(`Sample Name`, `Quantity Mean`, quant) %>%
  mutate(quant = ifelse(is.na(quant), Quantity, quant)) %>%
  #filter(!is.na(quant)) %>%
  mutate(quant_per_ul = quant/2)

plotA_2021 <- combined_data %>%
  left_join(., clean_meta, by = c("Sample Name" = "sample")) %>%
  filter(!is.na(tide) &
           distance != 100) %>%
  ggplot(aes(x = distance, y = log10(quant_per_ul), color = tide, shape = tide)) +
  geom_point(size = 2, alpha = 0.65) +
  geom_smooth(method = "lm", se = F) +
  #facet_grid(rows = vars(tide)) +
  theme_bw() +
   scale_shape_manual(values = c(16, 18)) +
  scale_color_manual(values = mycols) +
  labs(x = "Distance from pens (m)",
       y = "DNA copy number/uL (log10)",
       color = "Tide",
       shape = "Tide") +
  theme(
    axis.title.x = element_text(margin = margin(t = 5)),
    #axis.title.x = element_blank(),
    axis.title.y = element_text(margin = margin(r = 5)),
        legend.position = c(0.8, 0.8)
    #legend.background = element_rect(color = "gray50")
  )

plotA_2021
#ggsave("pdf_outputs/Amalga2021_qPCR_by_tide.png", width = 5, height = 4)
```


At incoming vs. outgoing tides, the concentration of chum salmon DNA at ~1000 m looks different.
Make a note of the LOD/LOQ in the figure legend, but include detections below the LOQ on the plot.

Make a plot of the CV for each distance:
```{r cv-by-distance-2021}
plotB_2021 <- combined_data %>%
  left_join(., clean_meta, by = c("Sample Name" = "sample")) %>%
  filter(!is.na(tide) & 
           !is.na(quant_per_ul) &
           distance != 100) %>%
  group_by(distance, tide) %>%
  mutate(sd = sd(log10(quant_per_ul))) %>%
  mutate(mean = mean(log10(quant_per_ul))) %>%
  dplyr::select(distance, sd, mean, tide) %>%
  unique() %>%
  mutate(cv = sd/mean) %>%
  ggplot(aes(x = distance, y = cv, color = tide, shape = tide)) +
  geom_point(size = 2, alpha = 0.85) +
  theme_bw() +
   scale_shape_manual(values = c(16, 18)) +
  scale_color_manual(values = mycols) +
  labs(x = "Distance from pens (m)",
       y = "CV",
       color = "Tide",
       shape = "Tide") +
  theme(
    axis.title.x = element_text(margin = margin(t = 5)),
    axis.title.y = element_text(margin = margin(r = 5)),
        #legend.position = c(0.8, 0.8)
    legend.position = "none"
  )

#ggsave("pdf_outputs/Amalga2021_qPCR_by_tideCV.png", width = 5, height = 4)

# combine plots A and B

plotA_2021 + plotB_2021 + plot_layout(ncol = 1, heights = c(2,1)) +
  plot_annotation(tag_levels = "A")

ggsave("pdf_outputs/Amalga2021_transect_and_CV.png", width = 5, height = 6)
```

Which samples are inhibited?
```{r inhibited-samples-plot}
fig2.a <- combined_data %>%
  left_join(., clean_meta, by = c("Sample Name" = "sample")) %>%
    filter(!is.na(tide) &
             distance != 100) %>%
  mutate(inhibited = ifelse(is.na(inhibited), "not inhibited", "inhibited")) %>%
  ggplot(aes(x = distance, y = log10(quant_per_ul))) +
  geom_point(aes(color = inhibited), alpha = 0.8) +
  facet_grid(rows = vars(tide), labeller = label_both) +
  theme_bw() +
  labs(x = "Distance from pens (m)",
       y = "DNA copy number/uL (log10)") +
  theme(
    axis.title.x = element_text(margin = margin(t = 5)),
    axis.title.y = element_text(margin = margin(r = 5)),
    legend.position = c(0.8, 0.9),
    #legend.background = element_rect(color = "gray50"),
    legend.text = element_text(size = 8),
    legend.title = element_blank(),
    legend.key.size = unit(0.4, "cm")
    #legend.spacing.y = unit(0.2, "cm")
  ) +
  scale_color_manual(values = c("darkblue", "darkgray"))

fig2.a
#ggsave("pdf_outputs/Amalga2021_qPCR_inhibition_by_tide.png", width = 6, height = 5)
```




## 2022 data



```{r read-in-list-of-plates-for-2022}
fp <- "../data/qpcr/2022/"

fp %>%
  list.files() %>%
  .[str_detect(., ".xlsx")] -> xlsx_names

# Load everything into the Global Environment
xlsx_names %>%
  purrr::map(function(file_name){ # iterate through each file name
  
  read_xlsx(paste0(fp, file_name), sheet = "Results", skip = 34)
  
}) -> df_list1 # Assign to a list

plates2022 <- bind_rows(df_list1, .id = "plate")

# multiple quantities by 10 because of 1:10 dilution factor
# and 2 ul of diluted extract added to each well
plates2022_10x <- plates2022 %>%
  mutate(quant = (Quantity*10)/2)

# list of samples
list2022 <- plates2022_10x %>%
  filter(Task =="UNKNOWN") %>%
  dplyr::select(`Sample Name`) %>%
  unique()

list2022
```

```{r}
# read in 2022 metadata 
meta2022 <- read_xlsx("../data/Amalga2022_cleanMetadata.xlsx")

meta2022

qpcr_meta2022_combined <- plates2022_10x %>%
  left_join(., meta2022, by = c("Sample Name" = "Extraction ID")) %>%
  filter(Task == "UNKNOWN" &
          !is.na(distance_from_pens)) 

qpcr_meta2022_combined %>%
  filter(`Target Name` == "Oke_COI" &
           transect == "perpendicular") %>%
  ggplot(aes(x = distance_from_pens, y = log(quant))) +
  geom_point(aes(color = time_of_day), alpha = 0.5) +
  facet_grid(cols = vars(transect), rows = vars(depth), space = "free", labeller = label_both) +
  theme_bw()


list2022 %>%
  left_join(., meta2022, by = c("Sample Name" = "Extraction ID")) %>%
  filter(Sample_Date == "2022_05_05")
  
```


```{r combine-data-from-both-years}
# 2021 data - including the inhibited samples rerun as dilutions
df2021 <- combined_data %>%
  filter(`Target Name` == "Oke_COI" &
           Task == "UNKNOWN") %>%
  dplyr::select(`Sample Name`, CT, quant_per_ul)
  
  
# 2022 data
df2022 <- plates2022_10x %>%
  filter(`Target Name` == "Oke_COI" &
           Task == "UNKNOWN") %>%
  rename(quant_per_ul = quant) %>%
  dplyr::select(`Sample Name`, CT, quant_per_ul)


# combine those
combo_df <- bind_rows(df2021, df2022) %>%
  rename(sample = `Sample Name`)
```


```{r make-metadata-consistent-btwn-years}
slim_m2021 <- clean_meta %>%
  mutate(time_of_day = location) %>%
  mutate(depth = 0) %>%
  mutate(transect = "perpendicular") %>%
  rename(distance_from_pens = distance) %>%
  dplyr::select(-location, -type, -dist_grp) %>%
  mutate(sample_date = "2021_05_10")

slim_m2022 <- meta2022 %>%
  rename(sample = `Extraction ID`, label = `Sample Code`, sample_date = Sample_Date) %>%
  dplyr::select(sample, label, distance_from_pens, time_of_day, depth, transect, sample_date) %>%
  mutate(tide = ifelse(time_of_day == "AM", "outgoing", "incoming")) 
  


metadata_both_years <- bind_rows(slim_m2021, slim_m2022) 
```

```{r combine-qpcr-data-and-metadata-for-both-years}
merged_df_both_years <- combo_df %>%
  left_join(., metadata_both_years) %>%
  filter(sample_date %in% c("2021_05_10", "2022_05_05")) %>%
  mutate(year = ifelse(sample_date == "2021_05_10", "2021", "2022")) %>%
  filter(distance_from_pens != 100) %>%
    mutate(log_dna_ul = log10(quant_per_ul))

merged_df_both_years %>%
  filter(year == "2022" & depth == 0)

merged_df_both_years %>%
  filter(year == "2022") %>%
  group_by(sample, tide, depth) %>%
  tally() %>%
  group_by(tide, depth) %>%
  tally()

merged_df_both_years %>%
  filter(year == 2021) %>%
  group_by(sample, tide) %>%
  tally() %>%
  group_by(tide) %>%
  tally()
```

Okay, that's kind of useful. Very little to talk about with the 2022 samples. There's one outlier for the incoming tide at 0m that I don't like - it would be nice to take a look at the replicates.


```{r}
merged_df_both_years %>%
  filter(sample_date == "2022_05_05") %>%
  mutate(detection = ifelse(CT == "Undetermined", "no", "yes")) %>%
  group_by(detection) %>%
  tally()

```
So many non-detections!

```{r}
merged_df_both_years %>%
  filter(sample_date == "2021_05_10") %>%
  mutate(detection = ifelse(CT == "Undetermined", "no", "yes")) %>%
  group_by(detection) %>%
  tally()

```

## Stats for distance & tide

Are data normally distributed?

```{r test-2021-data-for-normality}
for_mod2021 <- merged_df_both_years %>%
  filter(sample_date == "2021_05_10" &
           distance_from_pens != 100) # bec this is between the creek and the pens




ggdensity(for_mod2021$log_dna_ul)
ggqqplot(for_mod2021$log_dna_ul)

```

Those visual inspection methods make the data appear normally distributed, but it's good to test this explicitly.
```{r test-for-normality}
shapiro.test(for_mod2021$log_dna_ul)

```
Based on this, the data are not significantly different from normal.



```{r linear-regression-models}
# model 1: copy number and distance
mod1 <- lm(data = for_mod2021, log_dna_ul ~ distance_from_pens)

# model 2: distance and tide
mod2 <- lm(data = for_mod2021, log_dna_ul ~ distance_from_pens + tide)

# now compare the models with anova
anova(mod1, mod2)

summary(mod2)
```


```{r detections-over-distance}
fig2.d <- merged_df_both_years %>%
 filter(!is.na(distance_from_pens) &
           year == 2021 &
          distance_from_pens != 100) %>%
  group_by(tide, distance_from_pens) %>%
  mutate(detection = ifelse(is.na(quant_per_ul), 0, 1)) %>%
  summarise(n_detections = sum(detection)/9) %>%
  ggplot(aes(x = distance_from_pens, y = n_detections, fill = tide)) +
  geom_bar(stat = "identity") +
  facet_grid(rows = vars(tide), labeller = label_both) +
  theme_bw() +
  scale_shape_manual(values = c(1,2)) +
  scale_fill_manual(values = mycols) +
  labs(x = "Distance from pens (m)",
       y = "Proportion of detections") +
  theme(
    axis.title.x = element_text(margin = margin(t = 5)),
    axis.title.y = element_text(margin = margin(r = 5)),
    legend.position = "none"
  ) 

#ggsave("pdf_outputs/eDNA_detections2021.png", width = 6, height = 4)
```

```{r make-figure-2}
# combine the pieces from the other plots
plotA_2021 / (fig2.a | fig2.d) + plot_annotation(tag_levels = "A")

ggsave("pdf_outputs/Figure2_tri.png", width = 9, height = 8)

```




```{r dataframe-for-3d-plotting}
df_for_3d <- merged_df_both_years %>%
  filter(year == "2022") %>%
  mutate(distance_from_pens = ifelse(str_detect(transect, "_B"), -distance_from_pens, distance_from_pens)) %>%
  mutate(x_dist = ifelse(transect == "perpendicular", distance_from_pens, NA)) %>%
  mutate(y_dist = ifelse(str_detect(transect,"parallel"), distance_from_pens, NA)) %>%
  mutate(y_dist = ifelse(is.na(y_dist), 0, y_dist)) %>%
  mutate(x_dist = ifelse(is.na(x_dist), 1000, x_dist)) %>%
  mutate(log_quant = log10(quant_per_ul)) %>%
  mutate(depth = ifelse(depth != 0, -1*depth, depth))

```

```{r plots-for-depth-figure}
# I want to explicitly plot NAs on here to show where we sampled but didn't have data

df_for_3d %>%
  mutate(depth = -1*depth) %>%
  filter(!is.na(depth)) %>%
  ggplot(aes(x = x_dist, y = y_dist, size = log_quant, color = tide, shape = tide)) +
  geom_jitter(alpha = 0.5, width = 100, height = 100) +
  theme_bw() +
  facet_grid(rows = vars(depth), labeller = label_both) +
  labs(x = "Perpendicular transect (meters from pens)",
       y = "Parallel transect (SE-NW, meters from 1000 m midpoint)",
       color = "Tide",
       shape = "Tide",
       size = "DNA copies/uL\n(log10)") +
  theme(
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  ) +
  scale_shape_manual(values = c(16, 18), na.value = 1) +
  scale_color_manual(values = mycols) +
  guides(color = guide_legend(override.aes = list(size = 5)))

ggsave("pdf_outputs/depth_transects2022.png", width = 6, height = 7)

```

```{r explicitly-plot-NAs}
# I want to explicitly plot NAs on here to show where we sampled but didn't have data
tmp <- df_for_3d %>%
  mutate(depth = -1*depth) %>%
  filter(!is.na(depth))


### try a simple version
tmp %>%
  ggplot() +
  geom_jitter(aes(x = x_dist, y = y_dist, color = log_quant), size = 1.5, alpha = 0.9, width = 100, height = 70) +
  geom_point(data = tmp[is.na(tmp$log_quant),], aes(x_dist, y_dist), color = "gray", shape = 1, size = 3) +
  theme_bw() +
  facet_grid(rows = vars(depth), cols = vars(tide), labeller = label_both) +
  scale_color_gradient(na.value = NA) +
  #scale_shape_manual(na.value = 2) +
  labs(x = "Perpendicular transect (meters from pens)",
       y = "Parallel transect (SE-NW, meters from 1000 m midpoint)",
       color = "DNA copies/uL\n(log10)") +
   theme(
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  )

ggsave("pdf_outputs/depth_transects2022_v2.png", width = 6, height = 4)
```
Quick look at the dispersion across the parallel transect:
```{r}
df_for_3d %>%
  filter(#depth == 0 &
           x_dist == 1000 #&
           #!is.na(log_quant)
             ) %>%
  group_by(depth, tide) %>%
  tally()
  # ggplot(aes(x = y_dist, y = log_dna_ul, color = tide)) +
  # geom_point(size = 2, alpha = 0.5)

```




Relationship between qPCR detections and depth
```{r}
df_for_3d %>%
  mutate(depth = -1*depth) %>%
  filter(!is.na(depth) &
           transect == "perpendicular") %>%
  ggplot(aes(x = depth, y = log_quant, color = log_quant)) +
  geom_point() +
  facet_grid(rows = vars(tide), cols = vars(transect)) +
  scale_x_continuous(breaks = c(0, 5, 10)) +
  theme_bw() +
  theme(
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  ) +
    scale_color_gradient(na.value = "gray") +
  guides(color = guide_legend(override.aes = list(size = 5))) +
   labs(x = "Depth (m)",
       y = "DNA copies/uL\n(log10)",
       color = "Tide",
       shape = "Tide",
       size = "DNA copies/uL\n(log10)")

```
Is it worth fitting a linear model to this?

```{r perpendicular-depths}
dis22p <- df_for_3d %>%
  #mutate(depth = -1*depth) %>%
  filter(!is.na(depth) &
           transect == "perpendicular") %>%
  ggplot(aes(x = x_dist, y = depth, size = log_quant, color = tide, shape = tide)) +
  geom_jitter(alpha = 0.5, width = 70, height = 0.5) +
  facet_grid(rows = vars(tide), cols = vars(transect)) +
  #geom_smooth(method = "lm", se = F) +
  theme_bw() +
  theme(
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  ) +
  scale_shape_manual(values = c(16, 18)) +
  scale_color_manual(values = mycols) +
  scale_y_continuous(breaks = c(0, -5, -10)) +
   guides(color = guide_legend(override.aes = list(size = 5))) +
   labs(y = "Depth (m)",
       x = "Distance from pens (m)",
       color = "Tide",
       shape = "Tide",
       size = "DNA copies/uL\n(log10)")

ggsave("pdf_outputs/depth_by_distance.png", width = 6, height = 5)
```



```{r parallel-depths}
df_for_3d %>%
  #mutate(depth = -1*depth) %>%
  mutate(transect = ifelse(distance_from_pens == 1000 & transect == "perpendicular", "parallel", transect)) %>%
  mutate(y_dist = ifelse(distance_from_pens == 1000 & transect == "parallel", 0, y_dist)) %>%
  filter(!is.na(depth) &
           str_detect(transect, "parallel")) %>%
  ggplot(aes(x = y_dist, y = depth, size = log_quant, color = tide, shape = tide)) +
  geom_jitter(alpha = 0.5, width = 25, height = 0.3) +
  #facet_grid(rows = vars(tide), cols = vars(transect)) +
  #geom_smooth(method = "lm", se = F) +
  theme_bw() +
  theme(
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  ) +
  scale_shape_manual(values = c(16, 18)) +
  scale_color_manual(values = mycols) +
  scale_y_continuous(breaks = c(0, -5, -10)) +
   guides(color = guide_legend(override.aes = list(size = 5))) +
   labs(y = "Depth (m)",
       x = "Distance from pens (m)",
       color = "Tide",
       shape = "Tide",
       size = "DNA copies/uL\n(log10)")

ggsave("pdf_outputs/depth_by_distance_parallel.png", width = 6, height = 5)

```
% of detections in surface samples vs. at depth?





### Depth Stats - ANOVAs

1. Test whether there is statistically more DNA at the surface compared to the 5m and 10m depths
2. and then test whether incoming/outgoing tide makes a difference

```{r}
for_anova <- merged_df_both_years %>%
  filter(year == 2022 &
           !is.na(depth) &
           !is.na(quant_per_ul) &
           distance_from_pens == 0) %>%
  group_by(depth, tide) 

for_anova %>%
  summarise(sd = sd(log_dna_ul))
  
```
171 samples per depth from 2022, but the number of detections per depth is much smaller.

multiple comparisons:
every depth x tide compared to every other depth x tide

Do I need to fit a linear regression model first?

```{r nested-anova-depth-tide}
# set-up df
for_anova$depth <- as.factor(for_anova$depth)
for_anova$tide <- as.factor(for_anova$tide)

# form
# aov(response ~ factor A/ factor B)

depth.tide <- aov(for_anova$log_dna_ul ~ for_anova$tide / for_anova$depth, data = for_anova)
summary(depth.tide)

tukey.depth.tide <- TukeyHSD(depth.tide, conf.level = 0.95)
as.data.frame(tukey.depth.tide$`for_anova$tide:for_anova$depth`) %>%
  arrange(`p adj`) %>%
  filter(`p adj` < 0.05)
```
```{r}
plot(TukeyHSD(depth.tide, conf.level = 0.95), las = 2)
```



```{r boxplot-for-tide-and-depth}
d.boxp <- ggplot(for_anova, aes(x=factor(depth), y=log_dna_ul, fill=tide)) +
  geom_boxplot() +
  theme_minimal() +
  scale_fill_manual(values = c("coral3", "cadetblue4")) +
  theme_bw() +
  theme(
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
    #legend.position = c(0.85,0.85)
  ) +
   labs(x = "Depth (m)",
       y = "DNA copies/uL(log10)",
       fill = "Tide")

ggsave("pdf_outputs/depth_tide_boxplot.png", height = 4, width = 5)
```

```{r combine-depth-tide-anova-plots}
# make figure 4
dis22p + d.boxp + plot_annotation(tag_levels = "A") + plot_layout(ncol = 1, heights = c(3,2))

ggsave("pdf_outputs/depth_tide_anova_plot.png", width = 6, height = 7)
```



