---
title: "qPCR analysis"
output: html_notebook
---

9 December 2021


Organizing the initial qPCR results from the Amalga Harbor eDNA study in this updated R project.

Samples were taken at incoming and outgoing tide, every 80 m from the chum pens (DIPAC) perpendicular to the channel up to 2 km.

PM is outgoing tide; AM is incoming tide:
https://www.usharbors.com/harbor/alaska/juneau-ak/tides/?tide=2021-05#monthly-tide-chart

High tide is at 1:35 PM (14.8 ft)
Low tide is at 7:20 AM (-0.8 ft)

There are field replicates (3x per location per time [AM/PM]) and qPCR replicates (3x).


For the analysis, I want to check for outliers and then compare the correlation between distance from the chum pens and DNA copy number or quantity between the incoming and outgoing tides.




```{r load-packages}
library(tidyverse)
library(readxl)

```


```{r load-data}
# read in the extraction database metadata
meta <- read_xlsx("~/eDNA/extraction_database.xlsx", sheet = "metadata")

# read in the combined qPCR results
qpcr <- read_xlsx("~/eDNA/Amalga_study/qPCR_results_combined.xlsx")

```


### Metadata QC

```{r meta-qc}
# clean up the metadata format and variable names
frsh_meta <- meta %>%
  mutate(tide = ifelse(`Sample Code` == "AM", "incoming", "outgoing")) %>%
  mutate(tide = ifelse(is.na(`Sample Code`) | `Sample Code` == "Creek", NA, tide)) %>%
  rename(sample = `Extraction ID`, location = `Sample Code`, label = `Sample Label`, type = `Sample_type`) %>%
  # distance needs to be a numeric variable without meters attached
  mutate(dist_grp = ifelse(location %in% c("PM", "AM"), label, NA))

# strip off the meters on dist_grp so we can order them properly
frsh_meta$dist_grp <- gsub("[^0-9.-]", "", frsh_meta$dist_grp)

# make another variable that is numeric distance (not character)
clean_meta <- frsh_meta %>%
  mutate(distance = as.numeric(dist_grp)) %>%
  # remove some of the unneeded variables
  select(sample, location, label, type, tide, dist_grp, distance) %>%
  # add info to the creek samples
  mutate(label = ifelse(location == "Creek", "freshwater", label)) %>%
  mutate(label = ifelse(is.na(location), "blank", label))


```


### QC the qPCR data




