---
title: "2022 CTD data"
output: html_notebook
---

13 November 2022

Quick analysis of CTD data...


```{r load-libraries}
library(tidyr)
library(readr)
library(stringr)
library(lubridate)
library(hms)
library(dplyr)
library(ggplot2)
library(ggtext)

```


```{r read-in-csv-files}
# read in all files in the reheaded directory
dataFiles <- lapply(Sys.glob("../data/reheaded/*.csv"), read_csv)


# turn each of the data frames in that list into a single data frame
ctd.data <- do.call("rbind", dataFiles)


```


```{r data-clean-up}
# organize the headers, etc.
ctd.data2 <- ctd.data %>%
  rename(pressure_decibar = `Pressure (Decibar)`, depth_m = `Depth (Meter)`, temp_c = `Temperature (Celsius)`, conductivity = `Conductivity (MicroSiemens per Centimeter)`, sp_conduct = `Specific conductance (MicroSiemens per Centimeter)`, salinity = `Salinity (Practical Salinity Scale)`, sound_velocity = `Sound velocity (Meters per Second)`, density = `Density (Kilograms per Cubic Meter)`, file = `% File name`, lat = `% Start latitude`, long = `% Start longitude`, duration = `% Cast duration (Seconds)`, time = `% Cast time (local)`) %>%
  separate(file, into = c(NA, NA, "ctd_sample"), sep = "_") %>%
  select(ctd_sample, time, lat, long, depth_m, salinity, density, pressure_decibar, temp_c, conductivity, sp_conduct, sound_velocity, duration)


# add tide info
ctd.w.tides <- ctd.data2 %>%
  mutate(tide = ifelse(hour(hms::as_hms(time)) >12, "PM_incoming", "AM_outgoing")) %>%
    mutate(tide = ifelse(day(ymd_hms(time)) == "10", "PM_incoming", tide)) %>%
      mutate(tide = ifelse(day(ymd_hms(time)) == "25", "AM_incoming", tide))
  

ctd.w.tides %>%
  select(lat) %>%
  unique() %>%
  arrange(lat)

```


```{r}
ctd.depth <- ctd.w.tides %>%
  mutate(depth = depth_m *-1)
```

```{r}

```


Look at those data...

```{r}
library(ggrepel)
# maybe filter by date first
ctd.w.tides %>%
  filter(day(ymd_hms(time)) == "5") %>%
  select(ctd_sample, lat, long) %>%
  unique() %>%
  ggplot() +
  geom_point(aes(x = long, y = lat)) +
  geom_text_repel(aes(x = long, y = lat, label = ctd_sample))

```

Best way to connect ctd data (ctd_sample) to distance from pens via lat/lon?

```{r}
ctd.grp.df <- ctd.w.tides %>%
  #filter(day(ymd_hms(time)) == "25") %>%
  select(ctd_sample, lat, long, time) %>%
  unique() %>%
  arrange(time) %>%
  # round the lat/lon and then group by both lat/lon
  mutate(lat_r = round(lat, digits = 3)) %>%
  mutate(lon_r = round(long, digits = 3)) %>%
  group_by(lat_r, lon_r) %>%
  mutate(grp_id = cur_group_id()) %>% # add a group id for each lat/lon
  mutate(grp_id = ifelse(grp_id == 9, 7, grp_id)) %>%
  mutate(grp_id = ifelse(grp_id == 10, 8, grp_id)) %>%
  mutate(grp_id = ifelse(grp_id == 11, 9, grp_id)) %>%
  mutate(grp_id = ifelse(grp_id == 12, 10, grp_id))  

ctd.grp.df$grp_id <- as.character(ctd.grp.df$grp_id)
  
ctd.grp.df %>%
  ggplot() +
  geom_point(aes(x = long, y = lat, color = grp_id), size = 2) +
  geom_text_repel(aes(x = long, y = lat, label = ctd_sample)) +
  theme_minimal() +
  labs(
    x = "Longitude",
    y = "Latitude",
    title = "May 2022 - CTD samples"
  )
  
#ggsave("pdf_outputs/2022_Amalga_ctd_sample_locationsMay25.pdf", width = 8, height = 6)
```
But that only includes data from May 5, rather than all of the dates.

```{r}
tmp <- ctd.w.tides %>%
  #filter(day(ymd_hms(time)) == "5") %>%
  select(ctd_sample, lat, long, time) %>%
  unique() %>%
  arrange(time) %>%
  # round the lat/lon and then group by both lat/lon
  mutate(lat_r = round(lat, digits = 3)) %>%
  mutate(lon_r = round(long, digits = 2)) %>%
  group_by(lat_r, lon_r) %>%
   mutate(grp_id = cur_group_id()) %>%
  mutate(day = day(ymd_hms(time)))
  

tmp$grp_id <- as.character(tmp$grp_id)
tmp$day <- as.character(tmp$day)

tmp %>% # add a group id for each lat/lon
  mutate()
  filter(grp_id %in% c(11,6,7,9)) %>%
  ggplot() +
  geom_point(aes(x = long, y = lat, color = grp_id), size = 2) +
  #geom_text_repel(aes(x = long, y = lat, label = ctd_sample)) +
  theme_bw()

```


Thinking about these as spatial data points


Help from Pat...

```{r}
library(geosphere)

```

```{r}
ctd.simple <- ctd.grp.df %>%
  ungroup() %>%
  select(ctd_sample, lat, long, time)
```


```{r}
tmp <- geosphere::distm(x=cbind(ctd.grp.df$long, ctd.grp.df$lat),
                     fun = distGeo)

ctd.grp.df$ctd_sample

dist.matrix <- as.data.frame(tmp)
rownames <- ctd.grp.df$ctd_sample
rownames(dist.matrix) <- rownames
colnames(dist.matrix) <- rownames


# change that back into a df
dist.matrix %>%
  add_rownames(., var = "ctd_sample") %>%
  pivot_longer(2:43, names_to = "sample_num", values_to = "dist") %>% 
  filter(dist > 0, 
         dist < 100) %>% # remove comparisons from different stations
  arrange(dist) %>%
  # left_join(., ctd.simple) %>%
  # left_join(., ctd.simple, by = c("sample_num"= "ctd_sample"))
  mutate()

#ideally, now I would have a fixed point and add the distance to that point.
```












Now I can associate the ctd_sample and grp_id numbers with the distance from pens
```{r}
ctd.grp.w.data <- ctd.grp.df %>%
  left_join(., ctd.w.tides)

```

See next section with metadata...



```{r}
ctd.depth %>%
  filter(ctd_sample == "183334") %>%
  ggplot(aes(y = depth, x = salinity)) +
  geom_point() +
  theme_minimal() +
  theme(
    axis.title.x = element_text(margin = margin(t=10)),
     axis.title.y = element_text(margin = margin(t=10))
  ) +
  labs(y = "Depth (m)",
       title = "500 m NW from transect, AM")

```

```{r}
ctd.samples <- ctd.depth %>%
  select(ctd_sample) %>%
  unique()

x <- as.list(ctd.samples)
sample_nums <- x$ctd_sample
```


```{r}
plot.ctd <- function(sample, density){
  ctd.depth %>%
  filter(ctd_sample == sample) %>%
  ggplot(aes(y = depth, x = salinity)) +
  geom_point() +
  theme_minimal() +
  theme(
    axis.title.x = element_text(margin = margin(t=10)),
     axis.title.y = element_text(margin = margin(t=10))
  ) +
  labs(y = "Depth (m)",
       title = paste0("CTD sample ",sample)) 
}
```

```{r}
lapply(sample_nums, plot.ctd)
```






## Convert longitude to meters from chum pens

This will match the other plots in the series.

My logic previously was that the stations were sampled in the same order in the AM and PM. That changed a bit with the parallel transect.
Let's use the 5/5/2022 date as a first test case.

```{r}
meta <- read_csv("../data/2022_amalga_metadata_orderly.csv")

dist <- meta %>%
  filter(Sample_Date == "2022_05_05") %>%
  arrange(distance_from_pens) %>%
  select(distance_from_pens, transect, parallel_transect) %>%
  unique() %>%
  arrange(desc(transect), distance_from_pens)


```

groups:
1 = B3, -1000m parallel
2 = 2000m
3 = B2, -500m parallel
4 = background
5 = 1500m
6 = 1000m
7 = 500m
8 = T2, 500m parallel
9 = 0m
10 = T3, 1000m parallel



```{r}
# associate the grp id with the distance from pens
meta.grps <- meta %>%
  mutate(grp = ifelse(transect == "perpendicular" & distance_from_pens == 2000, 2, NA)) %>%
    mutate(grp = ifelse(transect == "perpendicular" & distance_from_pens == "1500", 5, grp)) %>%
    mutate(grp = ifelse(transect == "perpendicular" & distance_from_pens == "1000", 6, grp)) %>%
    mutate(grp = ifelse(transect == "perpendicular" & distance_from_pens == "500", 7, grp)) %>%
      mutate(grp = ifelse(transect == "perpendicular" & distance_from_pens == "0", 9, grp)) %>%
      mutate(grp = ifelse(transect != "perpendicular" & parallel_transect == "-1000", 1, grp)) %>%
        mutate(grp = ifelse(transect != "perpendicular" & parallel_transect == "-500", 3, grp)) %>%
          mutate(grp = ifelse(transect != "perpendicular" & parallel_transect == "500", 8, grp)) %>%
            mutate(grp = ifelse(transect != "perpendicular" & parallel_transect == "1000", 10, grp)) %>%
              mutate(grp = ifelse(transect != "perpendicular" & transect == "background", 4, grp)) 

meta.grps$grp <- as.character(meta.grps$grp)
```


```{r}
# combine the ctd data and the metadata
meta.ctd.df <- ctd.grp.w.data %>%
  left_join(., meta.grps, by = c("grp_id" = "grp")) %>%
  rename(sample_code = `Sample Code`)

```




```{r}
plot.ctd <- function(sample){
  ctd.grp.w.data %>%
  filter(ctd_sample == sample) %>%
  ggplot(aes(y = depth_m, x = salinity)) +
  geom_point() +
  theme_minimal() +
  theme(
    axis.title.x = element_text(margin = margin(t=10)),
     axis.title.y = element_text(margin = margin(t=10))
  ) +
  labs(y = "Depth (m)",
       title = paste0("CTD sample ",sample)) 
}
```

```{r}
lapply(sample_nums, plot.ctd)
```






```{r}
# each cast (ctd_sample) represents one of the 80m sampling intervals
ctd.w.dist <- ctd.w.tides %>%
  arrange(time) %>%
  select(time) %>%
  unique() %>%
  mutate(id = row_number()) %>%
  bind_cols(dist2x)

# bind it back together
ctd.w.tides.dist <- ctd.w.tides %>%
  left_join(., ctd.w.dist)
  
```
