---
title: "bathymap-for-amalga"
output: html_notebook
---

24 Feb 2022

```{r load-packages}
# library(tidyverse)
# library(ncdf4)
# library(raster)
```

```{r load-data}
# Let's see if this DEM from NOAA is useful
# juneau <- raster("JuneauBathy.tif")

# plot(juneau)
```
Filter out the land values by setting >=0

leaning on this: https://www.benjaminbell.co.uk/2019/08/bathymetric-maps-in-r-colour-palettes.html

To figure out a color scheme, need min and max
```{r}
# Calculate min and max values
# mi <- cellStats(juneau, stat="min")-100
# ma <- cellStats(juneau, stat="max")+100
```

```{r}
    # # Break points sequence for below sea level
    # s1 <- seq(from=mi, to=0, by=0 - mi / 50)
    # # Break points sequence for above sea level
    # s2 <- seq(from=0, to=ma, by=ma / 50)
    # 
    #     # Round sequence to nearest 100
    # s1 <- round(s1, -1)
    # s2 <- round(s2, -1)
    # 
    #     # Only show unique numbers
    # s1 <- unique(s1)
    # s2 <- unique(s2)
    # 
    # s1
    # 
    # s2
```
```{r}
    # Combine sequences and remove the first value from second sequence
    # s3 <- c(s1, s2[-1])
```


```{r}
    # Plot
    # plot(juneau, col=c(blue.col(50), terrain.colors(50)), breaks=s3)
```

```{r}
# crop <- extent(-134.8,-134.6,58.45,58.5)
# 
# extent(juneau)
# 
# june.zoom <- setExtent(juneau, crop)
# 
# plot(june.zoom, col=c(blue.col(50), terrain.colors(50)), breaks=s3)
```





## in ggplot

https://stackoverflow.com/questions/33227182/how-to-set-use-ggplot2-to-map-a-raster

```{r}
library(tidyverse)
library(ggplot2)
library(raster)
library(rasterVis)
library(rgdal)
library(grid)
library(scales)
library(viridis)  # better colors for everyone
library(ggthemes) # theme_map()
```

```{r}
datafold <- "../data/JuneauBathy.tif"
test <- raster(datafold) 
```


```{r}
test_spdf <- as(test, "SpatialPixelsDataFrame")
test_df <- as.data.frame(test_spdf)
colnames(test_df) <- c("value", "x", "y")

zoom.df <- test_df %>%
  filter(x > -134.95, x < -134.76, y > 58.4, y < 58.55)
  #filter(x > -135, x < -134, y > 58.2, y < 58.7) 

zoom.df.w.gray <- zoom.df %>%
  mutate(value = ifelse(value >=0, NA, value)) # set land to NA to shade it gray
```

```{r}
# minimum value below sea level
mi <- min(zoom.df$value)

# Break points sequence for below sea level
s1 <- seq(from=mi, to=0, by=0 - mi / 50)

depth.scale <- round(s1, 0)
depth.scale <- unique(depth.scale)

depth.scale
breaks = levels(depth.scale)[floor(seq(1, nlevels(depth.scale), length.out = 10))]

new.depth.scale <- depth.scale[seq(1, length(depth.scale), 3)]
```


```{r bathy-map-amalga}
zoomed.out <- ggplot() +  
  geom_tile(data=zoom.df.w.gray, aes(x=x, y=y, fill=value), alpha=0.8) + 
  scale_fill_steps2(low = "navyblue",
                    mid = "dodgerblue4",
  high = "lightsteelblue1",
  midpoint = -155,
  na.value = "grey50",
  breaks = new.depth.scale) +
  coord_equal() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 12),
        axis.title.x = element_text(size = 14, margin = margin(t=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10)),
        axis.line = element_line(color = "black"),
        legend.position="right") +
      theme(legend.text = element_text(size = 8)) + 
      theme(legend.key.height=unit(1.8, "cm")) +
      labs(fill = "Depth (m)",
         x = "Longitude (W)",
         y = "Latitude (N)")

zoomed.out

ggsave("pdf_outputs/amalgaBathy.pdf")
```



That's more zoomed out, but gives a better perspective...


Then here's more zoomed in on Amalga:


```{r}
more.zoom.df <- test_df %>%
  filter(x > -134.85, x < -134.77, y > 58.43, y < 58.51) %>%
  mutate(value = ifelse(value >=0, NA, value)) # set land to NA to shade it gray

zoomed.map <- ggplot() +  
  geom_tile(data=more.zoom.df, aes(x=x, y=y, fill=value), alpha=0.8) + 
  scale_fill_steps2(low = "navyblue",
                    mid = "dodgerblue4",
  high = "lightsteelblue1",
  midpoint = -97,
  #space = "Lab",
  na.value = "grey50",
  breaks = new.depth.scale) +
  coord_equal() +
  coord_equal() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 12),
        axis.title.x = element_text(size = 14, margin = margin(t=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10)),
        axis.line = element_line(color = "black"),
        legend.position="right") +
      theme(legend.text = element_text(size = 8)) + 
      theme(legend.key.height=unit(1.8, "cm")) +
      labs(fill = "Depth (m)",
         x = "Longitude (W)",
         y = "Latitude (N)")

zoomed.map

ggsave("amalgaBathyZoom.pdf")
```

Add the transect line to this map?




```{r}
# read in the data
# this dataframe is produced in the CTD analysis on the VM
# 08-ctd-cast-data.Rmd
ctd <- read_csv("../data/ctdDataframe.csv")

```


```{r}
# just one of the transects (not both AM and PM)
am <- ctd %>%
  filter(tide == "AM_incoming") %>%
  dplyr::select(lat, long, id, distance) %>%
  unique() %>%
  filter(id != 1)

am %>%
  ggplot(aes(x = long, y = lat, label = id)) +
  geom_text()
```

```{r}
zoomed.map +
  geom_point(data = am, aes(x = long, y = lat, label = id), size = 1, color = "firebrick2")

ggsave("pdf_outputs/amalgaTransectBathy.pdf", height = 6, width = 7)
```


Zoom more

```{r}
highest.res <- more.zoom.df %>%
  filter(x > -134.85, x < -134.77, y > 58.455, y < 58.51) %>%
  mutate(value = ifelse(value >=0, NA, value)) # set land to NA to shade it gray

highest.zoom.plot <- ggplot() +  
  geom_tile(data=highest.res, aes(x=x, y=y, fill=value)) + 
  scale_fill_steps2(low = "navyblue",
                    mid = "dodgerblue4",
  high = "lightsteelblue1",
  midpoint = -99,
  #space = "Lab",
  na.value = "grey50",
  breaks = new.depth.scale) +
  coord_equal() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 12),
        axis.title.x = element_text(size = 14, margin = margin(t=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10)),
        axis.line = element_line(color = "black"),
        legend.position="right") +
      theme(legend.text = element_text(size = 8)) + 
      theme(legend.key.height=unit(1, "cm")) +
      labs(fill = "Depth (m)",
         x = "Longitude (W)",
         y = "Latitude (N)")

```


```{r plot-for-manuscript}
transect2021 <- highest.zoom.plot +
  geom_point(data = am, aes(x = long, y = lat), size = 1, color = "red") +
  annotate(geom = "label", x = -134.795, y = 58.504, label = "Amalga Harbor", size = 4) +
  #annotate(geom = "label", x = -134.81, y = 58.494, label = "2021 transect", size = 3, fontface = "bold") +
  annotate(geom = "label", x = -134.78999, y = 58.496, label = "pens", size = 3, color = "black") +
    annotate(geom = "label", x = -134.808, y = 58.492, label = "1000m", size = 3, color = "black") +
      annotate(geom = "label", x = -134.825, y = 58.485, label = "2000m", size = 3, color = "black") +
  scale_x_continuous(limits = c(-134.83, -134.78), breaks = c(-134.82, -134.80, -134.78)) +
    scale_y_continuous(limits = c(58.47, 58.51)) 
  #labs(title = "2021 transect") +
  #scale_fill_continuous(breaks = c(0, -10, -25, -50, -75, -100, -125, -150, -200))
  
  
ggsave("pdf_outputs/amalgaTransectBathyZoom.pdf", width = 6, height = 5)

```
at 560 m, it's basically the edge of Amalga Harbor.


Add the transect to the zoomed out map:
```{r}
map2021_out <- zoomed.out +
  geom_point(data = am, aes(x = long, y = lat), size = 1, color = "black") +
  annotate(geom = "text", x = -134.786, y = 58.51, label = "Amalga Harbor", size = 4, color = "white", fontface = "bold") +
  annotate(geom = "label", x = -134.83, y = 58.494, label = "2021 transect", size = 3, fontface = "bold")
  
library(cowplot)

plot_grid(map2021_out, transect2021, rows = 2)
# if I choose to do that, I need to share the axes labels and legend.
```

With these maps, I should be able to combine the two transect years with cowplot or similar.


For the 2022 data, I need the equivalent from the CTD casts for the transect.

CTD data from 2022


















## NOT IN USE BELOW THIS LINE



## Bathy transect

```{r}
library(marmap)
```


```{r}
# load datasets
# 	data(nw.atlantic); as.bathy(nw.atlantic) -> atl
# 	data(nw.atlantic.coast)
# 
# # Example 1. get.transect(), without use of locator()
# 	get.transect(atl, -65, 43,-59,40) -> test ; plot(test[,3]~test[,2],type="l")
# 	get.transect(atl, -65, 43,-59,40, distance=TRUE) -> test ; plot(test[,4]~test[,3],type="l")
# 
# # Example 2. get.transect(), without use of locator(); pretty plot
# 	par(mfrow=c(2,1),mai=c(1.2, 1, 0.1, 0.1))
# 	plot(atl, deep=-6000, shallow=-10, step=1000, lwd=0.5, col="grey50",drawlabels=TRUE)
# 	lines(nw.atlantic.coast)
# 
# 	get.transect(atl, -75, 44,-46,32, loc=FALSE, dis=TRUE) -> test
# 	points(test$lon,test$lat,type="l",col="blue",lwd=2,lty=2)
# 	plotProfile(test)
# 
# # Example 3. get.transect(), with use of locator(); pretty plot
# ## Not run: 
# 	par(mfrow=c(2,1),mai=c(1.2, 1, 0.1, 0.1))
# 	plot(atl, deep=-6000, shallow=-10, step=1000, lwd=0.5, col="grey50",drawlabels=TRUE)
# 	lines(nw.atlantic.coast)
# 	
# 	get.transect(atl, loc=TRUE, dis=TRUE, col=2, lty=2) -> test
# 	plotProfile(test)
	
## End(Not run)
```


Very cool, so if I can get my data into that format, I should be able to do something similar.

```{r}
# -134.7925, -134.8213, 58.4946, 58.4848
# load dataset
juneau <- getNOAA.bathy(lon1 = -140, lon2 = -130,
lat1 = 60, lat2 = 55, resolution = 1)

plot(juneau)
summary(juneau)

```


```{r}
# Creating a custom palette of blues
blues <- c("lightsteelblue4", "lightsteelblue3",
"lightsteelblue2", "lightsteelblue1")
# Plotting the bathymetry with different colors for land and sea
plot(juneau, image = TRUE, land = TRUE, lwd = 0.1,
bpal = list(c(0, max(juneau), "grey"),
c(min(juneau),0,blues)))
# Making the coastline more visible
plot(juneau, deep = 0, shallow = 0, step = 0,
lwd = 0.4, add = TRUE)
```

zoom out even more to give perspective on location:

```{r}
juneau.out <- getNOAA.bathy(lon1 = -179, lon2 = -125,
lat1 = 75, lat2 = 48, resolution = 8)

plot(juneau.out, image = TRUE, land = TRUE, lwd = 0.1,
bpal = list(c(0, max(juneau.out), "grey"),
c(min(juneau.out),0,blues)))
# Making the coastline more visible
plot(juneau.out, deep = 0, shallow = 0, step = 0,
lwd = 0.4, add = TRUE)
```




That doesn't look as good as the DEM, but hopefully it can give us the bathy transect.

```{r}
# Example 1. get.transect(), without use of locator()
	get.transect(juneau, -134.8213, 58.4848, -134.7925, 58.4946) -> test ; plot(test[,3]~test[,2],type="l")
	get.transect(juneau, -134.8213, 58.4848, -134.7925, 58.4946, distance=TRUE) -> test ; plot(test[,4]~test[,3],type="l")

# Example 2. get.transect(), without use of locator(); pretty plot
	par(mfrow=c(2,1),mai=c(1.2, 1, 0.1, 0.1))
	plot(juneau, deep=-800, shallow=-10, step=1000, lwd=0.5, col="grey50",drawlabels=TRUE)
	#lines(nw.atlantic.coast)

	get.transect(juneau, -134.8213, 58.4848, -134.7925, 58.4946, loc=FALSE, dis=TRUE) -> test
	points(test$lon,test$lat,type="l",col="blue",lwd=2,lty=2)
	plotProfile(test)

# Example 3. get.transect(), with use of locator(); pretty plot
## Not run: 
	par(mfrow=c(2,1),mai=c(1.2, 1, 0.1, 0.1))
	plot(juneau, deep=-800, shallow=-10, step=1000, lwd=0.5, col="grey50",drawlabels=TRUE)
	lines(juneau)

	get.transect(juneau, loc=TRUE, dis=TRUE, col=2, lty=2) -> test
	plotProfile(test)
```
```{r}
bathy.df <- highest.res %>%
  rename(longitude = x, latitude = y, depth = value)

bathy.df %>%
  write_csv("amalgaBathyDf.csv")

read.bathy("amalgaBathyDF.csv", header = T)
```


```{r}
# Example 1. get.transect(), without use of locator()
	get.transect(highest.res, -134.8213, 58.4848, -134.7925, 58.4946) -> test ; plot(test[,3]~test[,2],type="l")
	get.transect(highest.res, -134.8213, 58.4848, -134.7925, 58.4946, distance=TRUE) -> test ; plot(test[,4]~test[,3],type="l")

# Example 2. get.transect(), without use of locator(); pretty plot
	par(mfrow=c(2,1),mai=c(1.2, 1, 0.1, 0.1))
	plot(test_spdf, deep=-800, shallow=-10, step=1000, lwd=0.5, col="grey50",drawlabels=TRUE)
	#lines(nw.atlantic.coast)

	get.transect(test_spdf, -134.8213, 58.4848, -134.7925, 58.4946, loc=FALSE, dis=TRUE) -> test
	points(test$lon,test$lat,type="l",col="blue",lwd=2,lty=2)
	plotProfile(test)

# Example 3. get.transect(), with use of locator(); pretty plot
## Not run: 
	par(mfrow=c(2,1),mai=c(1.2, 1, 0.1, 0.1))
	plot(test_spdf, deep=-800, shallow=-10, step=1000, lwd=0.5, col="grey50",drawlabels=TRUE)
	lines(test_spdf)

```






Also this:
https://stackoverflow.com/questions/47047623/projectraster-raster-projection-of-bathymetry-data-noaa-nc-in-the-pacific

